---
title: "PicantePez_Sep14"
author: "Carolina Paez"
date: "9/14/2020"
output: html_document
---
```{r}
library(picante)
library(permute)
library(lattice)
library(vegan)
library(tidyverse)
library(magrittr)
library(dplyr)
library(gdata)

comm <- read.csv("99_clustering_PICANTE.csv", as.is = TRUE, row.names = 1)
comm %<>% t() %>% as.data.frame()
comm <- decostand(comm, method="total")#hellinger: square root standarizaiton; hellinger: square root of method = "total" (Legendre & Gallagher 2001).
write.csv(comm, "sqr_picante_99.csv")

pogon.matrix <- data.matrix(comm, rownames.force = NA) #matrix

traits <- read.table(file = "traits.txt") #data.frame

metadata <-read.csv("meta_PICANTE.csv", header = TRUE, row.names = 1)
metadata %<>% rownames_to_column(var= "site")

phy <- read.tree(file = "99_PICANTE_2.newick") #phylo
phy

#Check for mismatches/missing species
combined <- match.phylo.comm(phy, comm)
phy <- combined$phy
phy
comm <- combined$comm

combined <- match.phylo.data(phy, traits)
phy <- combined$phy
traits <- combined$traits

all.equal(rownames(comm), rownames(metadata))

```
#Calculate Faith's PD
```{r}
comm.pd <- pd(comm, phy)
comm.pd
write.csv(comm.pd, "comm.pd_Apr23.csv")

#Test for PD differences among orientation
t.test(comm.pd$PD ~ metadata$orientation_NS)#nosignificant
```

#Compare PD and species richness
```{r}
plot(comm.pd$PD ~ comm.pd$SR, xlab = "Species richness", ylab = "Faith's PD")
```

#######################################################

Phylogenetic structure: MPD, MNTD, sesMPD, sesMNTD

Two approaches were used to calculate the observed phylogenetic distances among species in a community: Mean Pair Distance, which is more sensitive to phylogenetic patterns across the tree (dispersion), and Mean Nearest Taxon Distance, which is more sensitive to phylogenetic patterns closer to the tips (clustering) (Kembel et al., 2010).  

1. Unweighted Richness
1.1 MPD
1.2 MNTD

```{r}
library(ggplot2)
library(dplyr)
library(RColorBrewer)
#convert phylogeny to a distance matrix
phy.dist <- cophenetic(phy)
#1.1 MPD
comm.sesmpd <- ses.mpd(comm, phy.dist, null.model = "richness", abundance.weighted = FALSE, runs = 999, iterations = 1000)
comm.sesmpd
comm.sesmpd %<>% rownames_to_column(var= "site")
write.csv(comm.sesmpd, "richness_sesMDP.csv")

t.test(comm.sesmpd$mpd.obs.z ~ metadata$orientation_NS)


mycols <- 17
mycolors <- colorRampPalette(brewer.pal(8, "Dark2"))(mycols)

 
ggplot(comm.sesmpd, aes(site,mpd.obs.z, colour = site)) + 
  geom_point(size = 5, alpha = 0.55) + scale_color_manual(values = mycolors) +
  scale_y_continuous(name="ses.MPD", limits=c(-3, 3)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept=2, linetype="dashed", 
                color = "red", size=0.5) +
  geom_hline(yintercept=-2, linetype="dashed", 
                color = "red", size=0.5) +
   ggtitle("Richness ses.Mean Pair Distance") +
  theme(plot.title = element_text(hjust = 0.5))

#1.2 MNTD
comm.sesmntd <- ses.mntd(comm, phy.dist, null.model = "richness", abundance.weighted = FALSE, runs = 999)
comm.sesmntd
comm.sesmntd %<>% rownames_to_column(var= "site")
write.csv(comm.sesmntd, "richness_sesMNTD.csv")

ggplot(comm.sesmntd, aes(site,mntd.obs.z, colour = site)) + 
  geom_point(size = 5, alpha = 0.55) + scale_color_manual(values = mycolors) +
  scale_y_continuous(name="ses.MNTD", limits=c(-3, 3)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept=2, linetype="dashed", 
                color = "red", size=0.5) +
  geom_hline(yintercept=-2, linetype="dashed", 
                color = "red", size=0.5) +
   ggtitle("Richness ses.Mean Near Taxon Distance") +
  theme(plot.title = element_text(hjust = 0.5))


#A data frame of results for each community
##ntaxa Number of taxa in community
##mpd.obs Observed mpd in community
##mpd.rand.mean Mean mpd in null communities
##mpd.rand.sd Standard deviation of mpd in null communities
##mpd.obs.rank Rank of observed mpd vs. null communities
##mpd.obs.z Standardized effect size of mpd vs. null communities (= (mpd.obs - mpd.rand.mean)/ mpd.rand.sd, equivalent to -NRI)
##mpd.obs.p P-value (quantile) of observed mpd vs. null communities (= mpd.obs.rank / runs + 1)
##runs Number of randomizations
```


2. Unweigthed Independent swap
2.1 MPD
2.2 MNTD
```{r}
comm.sesmpd_IS <- ses.mpd(comm, phy.dist, null.model = "independentswap", abundance.weighted = FALSE, runs = 999)
comm.sesmpd_IS
comm.sesmpd_IS %<>% rownames_to_column(var= "site")
write.csv(comm.sesmpd_IS, "independentSwap_sesMDP.csv")

ggplot(comm.sesmpd_IS, aes(site,mpd.obs.z, colour = site)) + 
  geom_point(size = 5, alpha = 0.55) + scale_color_manual(values = mycolors) +
  scale_y_continuous(name="ses.MPD", limits=c(-3, 3)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept=2, linetype="dashed", 
                color = "red", size=0.5) +
  geom_hline(yintercept=-2, linetype="dashed", 
                color = "red", size=0.5) +
   ggtitle("Idependent Swap ses Mean Pair Distance") +
  theme(plot.title = element_text(hjust = 0.5))


#Calculate ses.mntd
comm.sesmntd_IS <- ses.mntd(comm, phy.dist, null.model = "independentswap", abundance.weighted = FALSE, runs = 999)
comm.sesmntd
comm.sesmntd_IS %<>% rownames_to_column(var= "site")
write.csv(comm.sesmntd, "independentSwap_sesMNTD.csv")


ggplot(comm.sesmntd, aes(site,mntd.obs.z, colour = site)) + 
  geom_point(size = 5, alpha = 0.55) + scale_color_manual(values = mycolors) +
  scale_y_continuous(name="ses.MNTD", limits=c(-3, 3)) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_hline(yintercept=2, linetype="dashed", 
                color = "red", size=0.5) +
  geom_hline(yintercept=-2, linetype="dashed", 
                color = "red", size=0.5) +
   ggtitle("Idependent Swap ses Mean Near Taxon Distance") +
  theme(plot.title = element_text(hjust = 0.5))

```


3. Weighted Rhichness: Distance among individuals
3.1 MPD
3.2 MNTD
```{r}
#3.1
comm.sesmpd_RW <- ses.mpd(comm, phy.dist, null.model = "richness", abundance.weighted = TRUE, runs = 999)
comm.sesmpd_RW
write.csv(comm.sesmpd_RW, "RW_sesMDP.csv")

#3.2
comm.sesmntd_RW <- ses.mntd(comm, phy.dist, null.model = "richness", abundance.weighted = TRUE, runs = 999)
comm.sesmntd_RW
write.csv(comm.sesmntd_RW, "RW_sesMNTD.csv")

```

4. Weighted Independent swap: Distance among individuals
4.1 MPD
4.2 MNTD
```{r}
#4.1
comm.sesmpd_ISW <- ses.mpd(comm, phy.dist, null.model = "independentswap", abundance.weighted = TRUE, runs = 999)
comm.sesmpd_ISW
write.csv(comm.sesmpd_ISW, "ISW_sesMDP.csv")

#4.2
comm.sesmntd_ISW <- ses.mntd(comm, phy.dist, null.model = "independentswap", abundance.weighted = TRUE, runs = 999)
comm.sesmntd_ISW
write.csv(comm.sesmntd, "ISW_sesMNTD.csv")
```


Tree with abundance heatmap
```{r}
# Plotting the tree
library(phangorn)
library(gdata)
library(ape)
library(ggtree)
library(tidyverse)
library(magrittr)
library(phytools)

data.set <- read.csv("sqr_picante_99.csv")
data.set[data.set==""] <- NA_character_  
  
t1 <- ggtree(phy)

t1 <- t1 + geom_tiplab(size=2, align=FALSE, linesize=.5)

t1

t1 <- ggtree(phy,layout = 'rectangular', branch.length = "none") %<+% data.set + geom_tippoint(aes(color=host),align=FALSE, hjust = -0.1, size = 5) +  scale_color_manual(values =c("#9590FF", "#00BF7D", "#E76BF3", "#00B0F6", "#D89000", "#969696"))
  

# To add the layers we need to remove the taxa names and add them as rownames for the data frame to be used
df.1 <- data.set
rownames(df.1) <- df.1[,1]
df.1 <- df.1[,-1]


library(ggnewscale)
library(wesanderson)
library(RColorBrewer)
library(viridis)


p1 <- t1 + new_scale_fill()
p1 <- gheatmap(p1, subset(df.1, select=c(3)), width=0.07, offset=0.076) + scale_fill_viridis_c(option = "plasma") + geom_treescale()
p1


p2 <- p1 + new_scale_fill()
p2 <- gheatmap(p2, subset(df.1, select=c(4)), width=0.07, offset=0.122) + scale_fill_viridis_c(option = "plasma")
p2

p3 <- p2 + new_scale_fill()
p3 <- gheatmap(p3, subset(df.1, select=c(5)), width=0.07, offset=0.168) + scale_fill_viridis_c(option = "plasma")
p3

p4 <- p3 + new_scale_fill()
p4 <- gheatmap(p4, subset(df.1, select=c(6)), width=0.07, offset=0.214) + scale_fill_viridis_c(option = "plasma")
p4

p5 <- p4 + new_scale_fill()
p5 <- gheatmap(p5, subset(df.1, select=c(7)), width=0.07, offset=0.260) + scale_fill_viridis_c(option = "plasma")
p5

p6 <- p5 + new_scale_fill()
p6 <- gheatmap(p6, subset(df.1, select=c(8)), width=0.07, offset=0.306) + scale_fill_viridis_c(option = "plasma")
p6

p7 <- p6 + new_scale_fill()
p7 <- gheatmap(p7, subset(df.1, select=c(9)), width=0.07, offset=0.352) + scale_fill_viridis_c(option = "plasma")
p7

p8 <- p7 + new_scale_fill()
p8 <- gheatmap(p8, subset(df.1, select=c(10)), width=0.07, offset=0.398) + scale_fill_viridis_c(option = "plasma")
p8

p9 <- p8 + new_scale_fill()
p9 <- gheatmap(p9, subset(df.1, select=c(11)), width=0.07, offset=0.444) + scale_fill_viridis_c(option = "plasma")
p9

p10 <- p9 + new_scale_fill()
p10 <- gheatmap(p10, subset(df.1, select=c(12)), width=0.07, offset=0.490) + scale_fill_viridis_c(option = "plasma")
p10

p11 <- p10 + new_scale_fill()
p11 <- gheatmap(p11, subset(df.1, select=c(13)), width=0.07, offset=0.536) + scale_fill_viridis_c(option = "plasma")
p11

p12 <- p11 + new_scale_fill()
p12 <- gheatmap(p12, subset(df.1, select=c(14)), width=0.07, offset=0.582) + scale_fill_viridis_c(option = "plasma")
p12 
```


#MANTEL TEST-PINE ONLY
##MANTEL distance decay

```{r}
library(vegan)
library(geosphere)
library(dplyr)
library(permute)
library(lattice)
library(tidyverse)
library(magrittr)
library(gdata)

#abundance data frame
abund <- read.csv("99_clustering_PICANTE.csv", as.is = TRUE, row.names = 1)
abund %<>% t() %>% as.data.frame()
abund <- decostand(abund, "hellinger") ##Hellinger-transform the species dataset

#Longitud and latitude
pogon.env <- read.csv("meta_PICANTE.csv", as.is = TRUE, row.names = 1)
geo <- data.frame(pogon.env$Long, pogon.env$Lat)

#abundance data frame - Bray-Curtis
dist.abund <- vegdist(abund, method = "bray")

#Geograohic data frame-haversine distance
d.geo <- distm(geo, fun = distHaversine)
dist.geo <- as.dist(d.geo)

#abundance vs geographic haversine distance
abund_geo  = mantel(dist.abund, dist.geo, method = "spearman", permutations = 9999, na.rm = TRUE)
abund_geo

#PLOT
#PLOTS
aa <- as.vector(dist.abund)
gg <- as.vector(dist.geo)

#new data frame with vectorized distance matrices
mat <- data.frame(aa, gg)

#abundance vs geographic distance plot
mm = ggplot(mat, aes(y = aa, x = gg/1000)) + 
    geom_point(size = 5, alpha = 0.5) + 
  xlim(0, 250) +
  labs(x = "Spatial distance (km)", y = "Bray-Curtis Dissimilarity") + 
    theme( axis.text.x = element_text(face = "bold",colour = "black", size = 12), 
        axis.text.y = element_text(face = "bold", size = 11, colour = "black"), 
        axis.title= element_text(face = "bold", size = 14, colour = "black"), 
        panel.background = element_blank(), 
        panel.border = element_rect(fill = NA, colour = "black"))
mm
```

#Correlation between Pinaceae richness (presence/absence matrix) and BC dissimilarity

```{r}
library(vegan)
library(geosphere)
library(dplyr)
library(permute)
library(lattice)
library(tidyverse)
library(magrittr)
library(gdata)

#abundance data frame all sites, all otus
abund <- read.csv("99_clustering.csv", as.is = TRUE, row.names = 1)
abund %<>% t() %>% as.data.frame()
abund <- decostand(abund, "hellinger") ##Hellinger-transform the species dataset

#Distance on binary data
pa_data <- read.csv("pinaceae_matrix.csv", as.is = TRUE, row.names = 1)
dist.pa_data <- vegdist(pa_data, binary = TRUE, method = "jaccard")

#abundance data frame - Bray-Curtis
dist.abund <- vegdist(abund, method = "bray")



#abundance vs geographic haversine distance
abund_pina  = mantel(dist.abund, dist.pa_data, method = "spearman", permutations = 9999, na.rm = TRUE)
abund_pina

#PLOT
#PLOTS
aa <- as.vector(dist.abund)
pp <- as.vector(dist.pa_data)

#new data frame with vectorized distance matrices
mat <- data.frame(aa, pp)

#abundance vs Pinaceae comm plot
mm = ggplot(mat, aes(y = aa, x = pp)) + 
    geom_point(size = 5, alpha = 0.5) + 
  labs(x = "Pinaceae community Jaccard", y = "Rhizopogon Bray-Curtis Dissimilarity") + theme( axis.text.x = element_text(face = "bold",colour = "black", size = 12), 
        axis.text.y = element_text(face = "bold", size = 11, colour = "black"), 
        axis.title= element_text(face = "bold", size = 14, colour = "black"), 
        panel.background = element_blank(), 
        panel.border = element_rect(fill = NA, colour = "black"))
mm
```

