---
title: "vsearch_97"
author: "Carolina Paez"
date: "4/6/2021"
output: html_document
---

#1. Loading data 99% clustering
```{r}
library(permute)
library(lattice)
library(vegan)
library(tidyverse)
library(magrittr)
library(dplyr)
library(gdata)
library(ggthemes)

#ALL data set; 17 sites and 15 otus.
otu_table <- read.csv("99_clustering.csv", as.is = TRUE, row.names = 1)
otu_table %<>% t() %>% as.data.frame()

taxo_table <- read.csv("Taxo_table_99.csv", as.is = TRUE, row.names = 1)

metadata <- read.csv("meta.csv", as.is = TRUE, row.names = 1)

#adding sampleid as a column
otu_table %<>% rownames_to_column(var= "site")
metadata %<>% rownames_to_column(var= "site")
site_joined_table <- left_join(metadata, otu_table, by="site")
write_csv(site_joined_table, "site_joined_table.csv")
```

#2. Calculating alpha diversity indices
```{r}
library(patchwork)
site.abundance.matrix <- site_joined_table[,13:33]
site.abundance.matrix<- decostand(site.abundance.matrix, method="total")#hellinger: square root standarizaiton; hellinger: square root of method = "total" (Legendre & Gallagher 2001).

site.indices <- site_joined_table[,c("island","site", "plant_community")]
site.indices$Richness <- rowSums(site.abundance.matrix>0)
site.indices$Shannon <- diversity(site.abundance.matrix) #shannon is default
site.indices$InvertedSimpson <- diversity(site.abundance.matrix, index = "invsimpson")

write_csv(site.indices, "site.indices_Apr6.csv")

#Plotting
forest_shan <- ggplot(site.indices, aes(plant_community, Shannon, fill=plant_community)) + 
  geom_boxplot() +  
  ggpubr::rotate_x_text(angle = 65) +
  theme(legend.position = "none") +
 scale_fill_manual(values=c("#40b8d0", "#b2d183"))
forest_shan 


forest_rich <- ggplot(site.indices, aes(plant_community, Richness, fill=plant_community)) + 
  geom_boxplot() +  
  ggpubr::rotate_x_text(angle = 65) +
  theme(legend.position = "none") +
 scale_fill_manual(values=c("#40b8d0", "#b2d183"))
forest_rich 


(forest_rich + forest_shan) 


#Shannon's diversity

#Multiple linear regression of differences in species richness based on range and forest stand
#ANOVA analysing diversity as a function of forest stand, range, and their interaction
shannon.model <- lm(Shannon ~ plant_community * island, data = site.indices)
anova(shannon.model)

t.shannon <- t.test(Shannon ~ plant_community, data = site.indices,
                        alternative = 'greater')
t.shannon

#ANOVA analysing Shannon's diversity as a function of forest stand
anova_result <-aov(Shannon ~ plant_community, site.indices)
summary(anova_result)


#Species richness

#Multiple linear regression of differences in species richness based on range and plant community
#ANOVA analysing diversity as a function of plant community, range, and their interaction
sr.model <- lm(Richness ~ plant_community * island, data = site.indices)
anova(sr.model)

t.rich <- t.test(Richness ~ plant_community, data = site.indices,
                        alternative = 'greater')
t.rich

#ANOVA analysing Richness as a function of forest stand
r.anova_result <-aov(Richness ~ plant_community, site.indices)
summary(r.anova_result)
```

#3. Barplot
```{r}
##RELATIVE ABUNDANCE BARPLOT
library(reshape)
#abundance table
otu <- read.csv("99_clustering_orderedbysubgenus.csv", as.is = TRUE, row.names = 1)
otu %<>% t() %>% as.data.frame()

##Relative abundance matrix from absolute abundance matrix
site.rel.matrix <- sweep(otu,1,rowSums(otu),"/")
write.csv(site.rel.matrix, "site.rel.matrix.csv")

otu.rel.org <- read.csv("site.rel.matrix.csv", as.is = TRUE, row.names = 1)
otu.rel.org %<>% rownames_to_column(var = 'site')

df_long <- melt(otu.rel.org, id = c("site"))
df_long$site <- factor(df_long$site, levels = unique(df_long$site))

library(RColorBrewer)
library(scales)
 
otu.colors <- c( "#6F2DA8", "#9590FF",
                 "#3E8914", "#3DA35D", "#AEF359",
                 "#8b0049", "#702e49",  "#E0115F","#FB607F", "#FF00FF", "#FBAED2", "#F0D8D8",
                 "#00A5FF", 
                  "#cc7722", "#c49102", "#FDA50f", "#E3B778", "#5d4f04", "#E6DBAC","#FFF200", "#BCC0BB")
     
mx <- ggplot(df_long, aes(x = site, fill = variable, y = value)) +
  geom_bar(stat = "identity", colour = "black", alpha = 0.76) +
  theme(axis.text.x = element_text(angle = 90, size = 14, colour = "black", vjust = 0.5, hjust = 1, face= "bold"), axis.title.y = element_text(size = 16, face = "bold"), legend.title = element_text(size = 16, face = "bold"), legend.text = element_text(size = 12, face = "bold", colour = "black"), axis.text.y = element_text(colour = "black", size = 12, face = "bold")) + 
    scale_y_continuous(expand = c(0,0)) + 
    labs(x = "", y = "Relative Abundance (%)", fill = "OTU") + 
    scale_fill_manual(values = otu.colors) + 
    theme_classic() +
    ggpubr::rotate_x_text(angle = 35)
  
mx
```

#4. We assessed completeness of sampling via species accumulation curves inferred in R and Vegan.
```{r}
##Calculating species accumulation curves

#Loading data
otu.data <- read.csv(paste0('site_joined_table.csv'),
                     as.is = T)
rownames(otu.data) <- otu.data$site


#Species accumulation curves: all data
#isolate community data
comm.matrix <- otu.data[13:length(otu.data)]

#create data frame
all.sa <- data.frame(community = rep(c('Overall','Pinus','Pseudotsuga'),
                           c(17,12,5)),
           Sites = NA,
           Richness = NA,
           sd = NA)

#Overall
overall.sa <- specaccum(comm.matrix)
all.sa[all.sa$community == 'Overall', 'Sites'] <- overall.sa$sites
all.sa[all.sa$community == 'Overall', 'Richness'] <- overall.sa$richness
all.sa[all.sa$community == 'Overall', 'sd'] <- overall.sa$sd

#Pinus
#isolate Pinus stands data

p.b <- rownames(otu.data[otu.data$plant_community == 'Pinus',])
p.b.data <- comm.matrix[rownames(comm.matrix) %in% p.b,]
pb.sa <- specaccum(p.b.data)
all.sa[all.sa$community == 'Pinus', 'Sites'] <- pb.sa$sites
all.sa[all.sa$community == 'Pinus', 'Richness'] <- pb.sa$richness
all.sa[all.sa$community == 'Pinus', 'sd'] <- pb.sa$sd


#Pseudotsuga
#isolate Pseudotsuga stands data
p.u <- rownames(otu.data[otu.data$plant_community == 'Pseudotsuga',])
p.u.data <- comm.matrix[rownames(comm.matrix) %in% p.u,]
pu.sa <- specaccum(p.u.data)
all.sa[all.sa$community == 'Pseudotsuga', 'Sites'] <- pu.sa$sites
all.sa[all.sa$community == 'Pseudotsuga', 'Richness'] <- pu.sa$richness
all.sa[all.sa$community == 'Pseudotsuga', 'sd'] <- pu.sa$sd


#plot
ggplot(all.sa, aes(x = Sites,
                   y = Richness,
                   color = community)) +
  geom_smooth(se = F) +
  theme_classic()


SAC <- ggplot(all.sa, aes(x = Sites,
                   y = Richness,
                   color = community)) +
  geom_point(data=all.sa, aes(x=Sites, y=Richness), size = 3) + 
  geom_line(data=all.sa, aes(x=Sites, y=Richness)) +
  geom_ribbon(data=all.sa ,aes(x=Sites,
                                   ymin=(Richness-2*sd),
                                   ymax=(Richness+2*sd)),
  alpha=0.1) +
  xlim(0, 25) +
  theme_bw() +
  expand_limits(y=c(0,25)) +
  ylab('OTUs') +
  xlab('Sites') +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  theme(axis.text = element_text(size=22, color = 'black'),
        axis.title = element_text(size = 28), legend.position = "none") +
  theme(axis.title.x = element_text(margin = margin(t = 30)),
        axis.title.y = element_text(margin = margin(r = 30))) +
  theme(legend.position = "bottom")

SAC
```

#5. ORDINATION
```{r}
library(magrittr)
library(tidyverse)
library(phyloseq)
library(vegan)
library(ggplot2)
library(ggthemes)
library(ggrepel)
library(viridis)
library(RColorBrewer)

# Data standarization: square root BRAY-CURTIS
OTUs <- read.csv("99_clustering.csv", as.is = TRUE, row.names = 1)
OTUs %<>% t() %>% as.data.frame()
OTUs_Sqrt<- decostand(OTUs, method="total")
set.seed(999)
min.r <- 0.1 #vector calculation, minimum r square; 0.1
min.p.val <- 0.05


Sqrt_nms <- metaMDS(OTUs_Sqrt, distance = "bray", trymax = 500) 
Sqrt_scores <- as.data.frame(scores(Sqrt_nms, display = "sites"))
Sqrt_scores %<>% rownames_to_column(var = 'site')
metadata %<>% rownames_to_column(var = 'site')
OTUs_Sqrt %<>% rownames_to_column(var = 'site')
taxo_table %<>% rownames_to_column(var = 'otu')

Sqrt_scores %<>% left_join(metadata, by = 'site') #ALL VARIABLES ASSOCIATED WITH YOUR SAMPLES
Sqrt_scores %<>% left_join(OTUs_Sqrt, by = 'site') #ALL VARIABLES WITH WHICH YOU ORDINATED YOUR SAMPLES

mycols <- 17
mycolors <- colorRampPalette(brewer.pal(8, "Dark2"))(mycols)
library(scales)
show_col(mycolors)

mycolors <- c( "#1B9E77", "#897932", 
                 "#C8611F", "#8D6B86", "#9B58A5",
                 "#DD2E8D", "#A66753",  "#70A61B", 
                 "#BBA90B") 

#SITE
ggplot(Sqrt_scores, aes(x = NMDS1, y = NMDS2,  color = island, shape = plant_community, label = site)) +
  geom_point(size = 6, alpha = 0.55) + 
  scale_color_manual(values = mycolors) +
  geom_text(size=2)                
  

#VECTORS using env.fit

Sqrt_nms.meta <- metadata %>% select(site, elevation_m, Lat, Long, ppt_mm, tmin_C, tmean_C, tmax_C, plant_community) #HERE'S YOUR CHANCE TO SPECIFY VECTORS FOR YOUR QUANTITATIVE ENV/SAMPLE DATA
Sqrt_nms.site_otu <- OTUs_Sqrt %>% select(site, "CAT1-RG-002",
"CAT1-RM-019",
"CAT2-RB-099",
"CAT2-RB-106",
"CAT2-RB-127",
"CAT2-RM-048",
"CHI1-CPP-2017-063",
"ELE2-RB-748",
"ELE2-RB-757",
"ELP1-RB-579",
"ELP1-RG-501",
"ELP1-RM-535",
"ELP2-RB-620",
"ELP2-RB-621",
"ELP2-RB-633",
"HUA1-CPP-2018-157",
"HUA2-RB-540",
"MAR2-RB-807",
"PIN1-RM-105",
"RIT1-RB-646",
"RIT1-RB-670") #HERE'S YOUR CHANCE TO SPECIFY VECTORS FROM YOUR MAIN RESPONSE (E.G. SPECIES) DATA

Sqrt_nms.combo <- left_join(Sqrt_nms.meta, Sqrt_nms.site_otu, by = 'site') #USE THIS IF YOUR SPECIFYING VECTORS FROM BOTH META AND MAIN VAR
Sqrt_nms.combo %<>% column_to_rownames("site")
Sqrt_nms.fit <- envfit(Sqrt_nms, Sqrt_nms.combo) #THIS IS WHERE VECTOR DATA ARE GENERATED--CHANGE SECOND TERM TO nms.meta, nms.var, or nms.combo

filt <- data.frame(r = Sqrt_nms.fit$vectors$r,pval = Sqrt_nms.fit$vectors$pvals) %>%
  rownames_to_column(var = "otu") %>%
  filter(r > min.r, pval < min.p.val) #SET R2 AND PVALUE FILTERS HERE FOR VECTORS

write_csv(filt, "99clustering__vectorsvalues.csv")

vector.scores <- as.data.frame(scores(Sqrt_nms.fit, display = "vectors"))
vector.scores %<>% rownames_to_column(var = 'otu')
vector.scores %<>% left_join(taxo_table, by = 'otu')
vector.scores %<>% filter(otu %in% filt$otu)
vector.scores %<>% column_to_rownames(var = 'otu')
vector.scores %<>% rownames_to_column (var = "vectors")

cols <- colorRampPalette(brewer.pal(8, "Dark2"))
pal <- cols(length(unique(metadata$Lat))) # SELECT DATA FRAME COLUMN HERE

Sqrt_plot <- ggplot(Sqrt_scores) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, color = island, shape = plant_community), size = 6, alpha = 0.55) + 
  geom_segment(data = vector.scores, #ADDS VECTORS TO YOUR ORDINATION; ADJUST THE SIZE/STYLE OF THE VECTORS HERE
               aes(x = 0, xend = NMDS1*.8, y = 0, yend = NMDS2*.8),
               arrow = arrow(length = unit(0.5, "cm")), colour = "black") +
  geom_label_repel(data = vector.scores, aes(x = NMDS1*1, y = NMDS2*1, label = vectors )) + 
                   #color = 'white',
                   #size = 2,
                   #segment.size = 1.5,
                   #segment.colour = 'gray',
                   #box.padding = 1) +
  scale_color_manual(values = mycolors) #USE THIS IF YOU WANT TO COLOR YOUR POINTS BY A CATEGORICAL VARIABLE
  #scale_fill_gradient2(midpoint = 3) + #OR USE THIS IF YOU WANT TO COLOR YOUR POINTS BY A CONTINUOUS VARIABLE
  #labs(color = "") + #USE THIS TO INDICATE WHAT VARIABLE YOU'RE COLORING BY (CURRENTLY BLANK)
  #ggtitle("") +
  #geom_text(mapping = aes(x = NMDS1, y = NMDS2, label=row.names(scores)), hjust=0, vjust=0) +
  #USE THE ABOVE IF YOU WANT TO LABEL YOUR POINTS WITH SAMPLE NAMES; THIS HELPS WITH OUTLIER INVESTIGATION
 
 
Sqrt_plot

```

#6. PERMANOVA
###Model 1: General model, all sites, all otus; this model would elucidate if there are general trends by island or by type of forest.

```{r}
############## PERMANOVA ###############
##1. Calculate Bray-Curtis Distance
Sqrt.BC.dist<-vegdist(OTUs_Sqrt, method='bray')

##2. PERMANOVA
m1_adonis <- adonis2(Sqrt.BC.dist ~ plant_community * island, by = "margin", data=metadata)
m1_adonis
```

#Model 2: Paired forest model: this model would elucidate if there are differences in the Rhizopogon communities when comparing these two forest stands within island

```{r}
#Loading data
otu_table <- read.csv("99_clustering_pineDF_only.csv", as.is = TRUE, row.names = 1)
OTUs_Sqrt <- decostand(otu_table, method="total")#hellinger: square root standarizaiton; hellinger: square root of method = "total" (Legendre & Gallagher 2001).a
OTUs_Sqrt %<>% t() %>% as.data.frame()
metadata <- read.csv("meta_pineDF_only.csv", as.is = TRUE, row.names = 1)
taxo_table <- read.csv("Taxo_table_99_pineDF_only.csv", as.is = TRUE, row.names = 1)


set.seed(999)
min.r <- 0.1 #vector calculation, minimum r square; 0.1
min.p.val <- 0.05

Sqrt_nms <- metaMDS(OTUs_Sqrt, distance = "bray", trymax = 500) # trymax ALLOWS YOU TO SET THE NUMBER OF ITERATIONS VEGAN GOES THROUGH
#'distance' CAN ALSO BE ADJUSTED AS NEEDED


Sqrt_scores <- as.data.frame(scores(Sqrt_nms, display = "sites"))#Options for display are sites, species
Sqrt_scores %<>% rownames_to_column(var = 'site')
metadata %<>% rownames_to_column(var = 'site')
OTUs_Sqrt %<>% rownames_to_column(var = 'site')
taxo_table %<>% rownames_to_column(var = 'otu')

Sqrt_scores %<>% left_join(metadata, by = 'site') #ALL VARIABLES ASSOCIATED WITH YOUR SAMPLES
Sqrt_scores %<>% left_join(OTUs_Sqrt, by = 'site') #ALL VARIABLES WITH WHICH YOU ORDINATED YOUR SAMPLES


mycolors <- c( "#1B9E77", "#897932", "#8D6B86", "#9B58A5", "#70A61B") 

#SITE
ggplot(Sqrt_scores, aes(x = NMDS1, y = NMDS2,  color = island, shape = plant_community, label = site)) +
  geom_point(size = 8, alpha = 0.7) + 
  scale_color_manual(values = mycolors) #+
  geom_text(size=2)                
  

#VECTORS using env.fit
Sqrt_nms.meta <- metadata %>% select(site, elevation_m, island, Lat, Long, ppt_mm, tmin_C, tmean_C, tmax_C, plant_community) #HERE'S YOUR CHANCE TO SPECIFY VECTORS FOR YOUR QUANTITATIVE ENV/SAMPLE DATA
Sqrt_nms.site_otu <- OTUs_Sqrt %>% select(site, "CAT1-RG-002",
"CAT1-RM-019",
"CAT2-RB-099",
"CAT2-RB-106",
"CAT2-RB-127",
"CAT2-RM-048",
"CHI1-CPP-2017-063",
"ELE2-RB-757",
"ELP1-RB-579",
"ELP1-RG-501",
"ELP1-RM-535",
"ELP2-RB-620",
"ELP2-RB-621",
"ELP2-RB-633",
"HUA1-CPP-2018-157",
"HUA2-RB-540",
"MAR2-RB-807",
"PIN1-RM-105",
"RIT1-RB-646",
"RIT1-RB-670") #HERE'S YOUR CHANCE TO SPECIFY VECTORS FROM YOUR MAIN RESPONSE (E.G. SPECIES) DATA


Sqrt_nms.combo <- left_join(Sqrt_nms.meta, Sqrt_nms.site_otu, by = 'site') #USE THIS IF YOUR SPECIFYING VECTORS FROM BOTH META AND MAIN VAR
Sqrt_nms.combo %<>% column_to_rownames("site")
Sqrt_nms.fit <- envfit(Sqrt_nms, Sqrt_nms.combo) #THIS IS WHERE VECTOR DATA ARE GENERATED--CHANGE SECOND TERM TO nms.meta, nms.var, or nms.combo

filt <- data.frame(r = Sqrt_nms.fit$vectors$r,pval = Sqrt_nms.fit$vectors$pvals) %>%
  rownames_to_column(var = "otu") %>%
  filter(r > min.r, pval < min.p.val) #SET R2 AND PVALUE FILTERS HERE FOR VECTORS

write_csv(filt, "model2_vector_values.csv")

vector.scores <- as.data.frame(scores(Sqrt_nms.fit, display = "vectors"))
vector.scores %<>% rownames_to_column(var = 'otu')
vector.scores %<>% left_join(taxo_table, by = 'otu')
vector.scores %<>% filter(otu %in% filt$otu)
vector.scores %<>% column_to_rownames(var = 'otu')
vector.scores %<>% rownames_to_column (var = "vectors")

cols <- mycolors <- c( "#1B9E77", "#897932", "#8D6B86", "#9B58A5", "#70A61B") 
pal <- cols(length(unique(metadata$Lat))) # SELECT DATA FRAME COLUMN HERE

Sqrt_plot <- ggplot(Sqrt_scores) +
  geom_point(mapping = aes(x = NMDS1, y = NMDS2, color = island, shape = plant_community), size = 8, alpha = 0.7) + 
  geom_segment(data = vector.scores, #ADDS VECTORS TO YOUR ORDINATION; ADJUST THE SIZE/STYLE OF THE VECTORS HERE
               aes(x = 0, xend = NMDS1*0.3, y = 0, yend = NMDS2*0.3),
               arrow = arrow(length = unit(0.3, "cm")), colour = "black") +
  geom_label_repel(data = vector.scores, aes(x = NMDS1*0.3, y = NMDS2*0.3, label = vectors )) + scale_color_manual(values = mycolors) 
 
Sqrt_plot

############## PERMANOVA ###############
##1. Calculate Bray-Curtis Distance
Sqrt.BC.dist<-vegdist(OTUs_Sqrt, method='bray')

##2. PERMANOVA
m1_adonis2 <- adonis2(Sqrt.BC.dist ~ plant_community + island, by = "margin", data=metadata)
m1_adonis2

m2_adonis <- adonis(Sqrt.BC.dist ~ plant_community, strata = metadata$island, data = metadata)
m2_adonis

m2.2_adonis <- adonis(Sqrt.BC.dist ~ island/plant_community, strata = metadata$island, data = metadata)
m2.2_adonis

m3_adonis2 <- adonis2(Sqrt.BC.dist ~ island, data = metadata)
m3_adonis2 
```







